Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-04-05 journal-draft/bare_jrnl_comsoc.tex
finished 6, and randomly proofread to shorten the pages, now 14 pages!

%% Packages
%\usepackage{tikz}
%\usetikzlibrary{shapes,arrows}
%\usetikzlibrary{positioning}
\usepackage{graphicx}
\usepackage{kantlipsum} %<- For dummy text
\usepackage{mwe} %<- For dummy images
\usepackage{multirow}
\usepackage{multicol}
%\usepackage{colortbl}
%\usepackage{todonotes}	
\newcounter{todocounter}
\newcommand{\todonum}[2][]
{\stepcounter{todocounter}\todo[inline,backgroundcolor=blue!10!white, #1]{\thetodocounter: #2}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Cluster structure in cognitive radio networks facilitates cooperative spectrum sensing, routing and other functionalities.
The availability of unlicensed channels which are available for every member in a cluster decides the survival of that cluster from licensed users' influence.
Thus in order to be robust against licensed users, there should be more unlicensed channels in the clusters.
In the process of forming clusters, every secondary user needs to decide with whom to form a cluster, or which cluster to join.
Congestion game model is adopted to analyse this process, which not only contributes the algorithm design directly, but also provides guarantee of convergence into Nash Equilibrium and convergence speed.
%
Our proposed distributed clustering scheme outperforms the comparison scheme in terms of robustness against primary users, convergence speed and volume of control messages.
Furthermore, the proposed clustering solution is versatile to fulfil other requirements such like fast convergence and cluster size control.
Besides, we prove the clustering problem to be NP-hard, and also propose the centralized solution.
The extensive simulation supports our claims.
\end{abstract}

@@

\section{Introduction}
\label{intro}
%CR concept
\todo[inline]{todo: polish section 6}
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Unlicensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, unlicensed users can only access the licensed spectrum after validating the channel is unoccupied by licensed users.
This refers to the process of sensing a particular channel and verifying (with a previously specified probability of error) that it is not used by a primary user currently.
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), and the CR users are known as secondary users and constitute the cognitive radio networks (CRN).

@@

Secondary users conduct spectrum sensing independently and sequentially on all licensed channels.
The sensing duration and frequency on one channel is a research topic~\cite{sensing_survey_2009}, and we assume that every node can detect the presence of primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
%In our system model, we assume that a secondary node $n$ is able to detect for each channel $c$ whether any primary user utilizes $c$.
We denote $K_i \subseteq \mathcal{K}$ as the set of available channels for $i$.

%One dedicated control channel is assumed to be available for all the secondary nodes to exchange control messages with neighbours in the process of cluster formation.
%Note that the assumption of dedicated control channel is to simplify the discussion so that we can focus on the kernel of this paper, the robustness of clusters.
%Actually, the control messages involved in the clustering process can be conveyed on available licensed channels through rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}. 
%Over the control channel, secondary users exchange their spectrum sensing results $V_{i}$ with one hop neighbours. 
%Secondary users have the same transmission range on both licensed and control channel, and the communication links are reciprocal.

We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
Both primary users and CR users have fixed transmission ranges respectively, and the all the channels are regarded to be identical in terms of signal propagation.
If a CR node locates within the transmission range of primary user $p$, that CR node is not allowed to use the channel $k(p)$.

We assume that in addition to the licensed channels, there is one dedicated control channel.
This control channel could be in ISM band or other reserved spectrum which is exclusively used for transmitting control messages.
Actually, the control messages involved in the clustering process can be transmitted on available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to any $i' \in \text{Nb}(i)$.
It is available for any secondary node $i$ to exchange control messages with any other node in its proximity (or neighborhood) $\text{Nb}(i)$ during the cluster formation phase.
$\text{Nb}(i)$ is simply defined as the set of nodes located within the transmission range of $i$.

@@

----------- No modifications from here on ... -------------

@@

\section{Distributed Coordination Framework: Clustering Algorithm}
\label{ross}

In this section, we introduce the distributed clustering scheme ROSS which leads to robust clusters against primary users' influence.
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters on the basis of the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

@@

\item \textit{Social connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It is the number of common channels available to$i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbours.
\end{itemize}
Individual connectivity degree $d_i$ and social connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where each node's connectivity vector is calculated and shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph and the connectivity vector $(d_i, g_i)$ at each node. The available channels sensed by each CR node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. The dashed edge indicates the end nodes are within each other's transmission range. The number of common channels between the two nodes is shown by the edge.}
	\label{fig1}
\end{figure}

@@

\subsubsection{Determining Cluster Heads and Form the Initial Clusters}
In this phase, each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than any neighbors except for those which have already become cluster heads (the appearance of cluster heads will be explained in Section~...), then node $i$ becomes clusters head.
%In other words, CR node $i$ becomes cluster head if $d_i > d_k, \forall k\in \texttt{Nb}_i\setminus CHs$ ($CHs$ donate the cluster heads existing in $\texttt{Nb}_i$).
If there is another CR node $j$ in its neighborhood which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\text{CHs}\cup i\}$, then the node out of $\{i, j\}$ with higher social connectivity degree becomes cluster head. The other nodes become a member of that cluster. 
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster's head.
%
The node which becomes cluster head broadcasts a message of its eligibility of being cluster head to notify its neighbours, and claims its neighbourhood as its cluster.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.

@@

\subsubsection{Guarantee the Availability of Common Control Channel}
%After deciding itself being cluster head, CR node broadcasts to notify its neighbours on the dedicated control channel, meanwhile, $i$'s initial cluster is formed immediately, which is $i$'s neighbourhood except for those nodes which have become cluster heads, \ie $C_i=(\text{Nb}_i\setminus \text{CHs})\cup i$.
After executing phase I of ROSS, there are some secondary users which become cluster heads.
The cluster head and its neighbourhood (except for the CHs) become a cluster.
It is possible that there is no CCCs for some formed clusters, and we solve this problem with the following method.

@@

In the following we illustrate the pressing necessity to control the cluster size when CRN becomes dense via both theoretical analysis and simulation.
Assuming the secondary and primary users are evenly distributed and primary users occupy the licensed channels randomly, then both CR nodes density and channel availability in the CRN can be seen to be spatially homogeneous.
%Based on Algorithm~\ref{alg0}, cluster heads are the CR nodes which possess the biggest individual connectivity degrees in their neighborhood respectively, and they are surrounded by CR nodes.
%In contrast, CR nodes residing on edge are unlikely to become cluster heads as their neighbourhoods are only half the nodes locating away from edge.
The formed clusters are the neighbourhoods of cluster heads, and the neighbourhood is decided by the transmission range and network density.
%
We consider a cluster $C(i)$ where $i$ is CH in a dense CRN. 
When we don't consider the CHs which could appear within $i$'s neighbourhood in the procedure of guaranteeing CCCs, according to Algorithm~\ref{alg0}, the nearest cluster heads could locate just outside node $i$'s transmission range.
An instance of this situation is shown in Figure~\ref{clusters_denseNetwork}.
%
In the figure, black dots represent cluster heads, the circles denotes the transmission ranges of cluster heads.
Cluster members are not shown in the figure.
%Circles represent the transmission range of cluster head, within which CR nodes are absorbed in cluster.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\linewidth]{clusters_denseNetwork_2.pdf}
  \caption{Clusters formation in extremely dense CRN. Black dots are cluster heads, other cluster members are not drawn.}
  \label{clusters_denseNetwork}
\end{figure}
Let $l$ be the length of side of simulation plan square, and $r$ be CR's transmission radius.
Based on the aforementioned analysis and geometry illustration as shown in Figure~\ref{clusters_denseNetwork}, we give an estimate on the maximum number of generated clusters, which is the product of the number of cluster heads in one row and that number in one line, $l/r * l/r = l^2/r^2$.
%Now we verify the estimation with simulation.
%We distribute CR and primary users randomly on a square plain, $r$=10 and $l$=50.
%Network density is increased by adding more CR users.
%We now have a look at how does the network density affect the cluster size when the transmission range is constant.
%This implicates when the cluster size is decided by the density of the network.
%As to SOC, the membership of one cluster is decided after a complex process, and the cluster size is roughly the same with one neighborhood.xxxx
%We can see from the example that although two neighbouring clusters can overlap greatly with each other, no cluster head will be covered by other clusters.
Given $r$=10 and $l$=50, the number of formed clusters is shown in Figure~\ref{number_clusters_scale}.
With the increase of CR users in the network, network density increases linearly (the Y axis label is the average number of neighbours), and the number of formed clusters also increases and approaches to the the upper bound of 25 which complies with the estimation.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\linewidth]{number_clusters_upperBound.pdf}
  \caption{The correlation between the number of formed clusters and network density. Note that the number of neighbours denotes the network density. Simulation is run for 50 times and the confidence interval is 95\%.}
  \label{number_clusters_scale}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.6\linewidth]{figure2.pdf}
  \caption{Clusters formation after the first phase of ROSS. There are some nodes being debatable nodes, \ie belonging to more than one cluster.}
  \label{fig2}
\end{figure}

\subsection{Phase II - Membership Clarification}
%\subsubsection*{Problem Description}
After applying phase I of ROSS to the example in Figure~\ref{fig1}, the resulted clusters are shown in Figure~\ref{fig2}.
We notice nodes $A, B, D$ are included in more than one cluster. 
We refer these nodes as \textit{debatable nodes} as their cluster affiliations are not clear, and the clusters which include the debatable node $i$ are called \textit{claiming clusters} of node $i$, and are denoted as $S_i$.  
Actually, debatable nodes extensively exist in CRN with larger scale.
Debatable nodes should be exclusively associated with only one cluster and be removed from the other claiming clusters, this procedure is called \textit{cluster membership clarification}.
We will introduce the solution for cluster membership clarification in the following.

@@

This procedure raises a concern that the debatable nodes may never stop changing their affiliations, as debatable nodes' choices seem to be dependent on each other, and the infinite chain effect never ceases.
%  update their choices based on other debatable nodes' choices, and this process 
For example, assuming one debatable node $i$ locates in cluster $C\in S_i$, and $C$ has more than one debatable node except for $i$.
Assuming node $i$ makes decision to stay in the claiming cluster $C$, afterwards one another debatable nodes $j$ decides its affiliation, and there is $j\in C\in S_i$.
When $j$ leaves cluster $C$, which decrease cluster $C$'s cluster size, and could possibly trigger node $i$ to alter its previous decision to leave $C$, as $C$'s size is smaller now and leaving it may result in more increase of CCCs in $S_i$.
At this point of time, debatable node $j$ may rejoin cluster $C$ duo to the changes in $S_j$, then both node $i$ and $j$ are facing the \textit{same}\footnote{Actually it is not totally same as before, as there are some new changes within $S_j$.} situation again.
%To maximize the increase of common channel in clusters in $S_j$, $j$ either stays in $C$, or leaves cluster $C$ and stay another cluster in $S_j$.
%$j$'s decision changes $C$'s membership, which will affect $i$'s decision.
%Assume $i$ makes decision before $j$ and chooses to stay in $C$, as which brings the most common channels in $i$' claiming clusters, and then it is $j$' turn to choose cluster.
%If $j$ leaves $C$, the smaller cluster $C$ will possibly make $i$ to leave it and join another cluster.
%and the choice of $j$ to stay in $C$ or not possibly changes $C$'s membership and  which potentially further triggers node $i$ to alter its previous decision. 
Thence, we must answer this concern raised when implementing ROSS-DGA.
%, and if it converges, how good such a distributed scheme performs. 
In the following we show that the process of membership clarification can be formulated into a singleton congestion game, and a equilibrium is reached after a finite number of best response updates made by the debatable nodes.

@@

To formulate the debatable nodes' membership clarification into the desired congestion game, we observe this process from a different (or opposite) perspective. 
From the new perspective, the debatable nodes are regarded to be isolated and don't belong to any cluster, in other words, their claiming clusters become clusters which are beside them. 
Now for the debatable nodes, the previous problem of deciding which clusters to leave becomes a new problem that which cluster to join.
In the new problem, debatable node $i$ chooses one cluster $C$ out of $S_i$ to join if the decrease of CCCs in cluster $C$ is the smallest in $S_i$, and the decrease of CCCs in cluster $C$ is $\sum_{C\in S_i}\Delta\vert K(C) \vert=\sum_{C\in S_i}({\vert K(C) \vert-\vert K(C\cup i) \vert})$.
The interaction between the debatable nodes and the claiming clusters is shown in Figure~\ref{debatable_nodes_claiming_cluster}.
%The concern on convergence appears again as we have discussed in the previous subsubsection.
We give proof on convergence under game theoretic framework.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.25\linewidth]{singletongame_png.png}
  \caption{Debatable nodes and claiming clusters}
  \label{debatable_nodes_claiming_cluster}
\end{figure}

@@

%\todo[inline]{check}
The centralized clustering scheme aims to form clusters with desired sizes, meanwhile the total number of common control channels of all clusters is maximized.
In the following, we refer this problem as \textit{centralized clustering}, and give the formal problem definition.

@@

The objective is a sum of two parts, the first part is the sum of products of cluster size and the corresponding number of CCCs.
The first part is the only metric adopted by the scheme SOC~\cite{Lazos09}.
The second part is the \textit{punishment} for choosing the clusters whose sizes are not $\delta$.
In fact, the second part is particularly designed to eliminates the drawbacks of SOC, \ie SOC produces a large number of singleton clusters and a few very large clusters which access affluent unlicensed spectrum.
In practical computation, we minimize the opposite of the first part, then the punishment is a positive value.

The first constraint restricts each node $j$ to reside in exactly one cluster.
In the second constraint, $w_i$ is an auxiliary binary variable, which denotes whether cluster $C_i$ is chosen by the solution, in particular, 
$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th legitimate cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th legitimate cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
The second constraint regulates that when the $i$th legitimate cluster $C_i$ is chosen, then  there are $|C_i|$ elements in $i$th row in matrix $X$, which are 1.
%
Now we explain how does the mechanism of the punishment in the objective work. 
The parameter $p$ is defined as follows,
$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$} \\
\dots
\end{array} \right.
$$
where $\alpha_i>0$ and increases when $|C_i|$ diverges from $\delta$.
Because of $w_i$, any chosen cluster ($w=0$) brings certain \textit{punishment}.
When the chosen cluster's size is desired size $\delta$, the punishment is zero.
In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
In particular, when $w_i=0$ and $|C_i|=1$, the punishment is the most severe.
This design doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
Choice of $\alpha_i$ affects the resultant clusters.

@@

\item \textit{Cluster sizes.}
Specific clusters size is pursued in many applications due to energy preservation and the system design ~\cite{clustering_globecom11}.
We will present the distribution of CRs residing in the formed clusters, and the number of generated clusters through multiple simulations.

>>>>

%% Packages
%\usepackage{tikz}
%\usetikzlibrary{shapes,arrows}
%\usetikzlibrary{positioning}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{kantlipsum} %<- For dummy text
\usepackage{mwe} %<- For dummy images
\usepackage{multirow}
\usepackage{multicol}
%\usepackage{colortbl}
%\usepackage{todonotes}	
\newcounter{todocounter}
\newcommand{\todonum}[2][]
{\stepcounter{todocounter}\todo[inline,backgroundcolor=blue!10!white, #1]{\thetodocounter: #2}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Cluster structure in cognitive radio networks facilitates cooperative spectrum sensing, routing and other functionalities.
The availability of unlicensed channels which are available for every member in a cluster decides the robustness of that cluster against the licensed users' influence.
Thus there should be more unlicensed channels in the clusters.
In the process of forming clusters, every secondary user decides with whom to form a cluster, or which cluster to join.
Congestion game model is adopted to analyse this process, which not only contributes the algorithm design directly, but also provides guarantee of convergence into Nash Equilibrium and convergence speed.
%
Our proposed distributed clustering scheme outperforms the comparison scheme in terms of cluster robustness against the primary users, convergence speed and volume of control messages.
Furthermore, the proposed clustering solution is versatile to fulfil other requirements such like fast convergence and cluster size control.
Besides, we prove the clustering problem is NP-hard, and also propose the centralized solution.
The extensive simulation supports our claims.
\end{abstract}

@@

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Unlicensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, unlicensed users can only access the licensed spectrum after validating the channel is unoccupied by licensed users.
This refers to the process of sensing a particular channel and verifying (with a previously specified probability of error) it is not used by a primary user currently.
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), and the CR users are known as secondary users and constitute the cognitive radio networks (CRN).

@@

Secondary users conduct spectrum sensing independently and sequentially on all licensed channels.
%The sensing duration and frequency on one channel is a research topic~\cite{sensing_survey_2009}, and 
We assume that every node can detect the presence of primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
%In our system model, we assume that a secondary node $n$ is able to detect for each channel $c$ whether any primary user utilizes $c$.
We denote $K_i \subseteq \mathcal{K}$ as the set of available channels for $i$.
%
%One dedicated control channel is assumed to be available for all the secondary nodes to exchange control messages with neighbours in the process of cluster formation.
%Note that the assumption of dedicated control channel is to simplify the discussion so that we can focus on the kernel of this paper, the robustness of clusters.
%Actually, the control messages involved in the clustering process can be conveyed on available licensed channels through rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}. 
%Over the control channel, secondary users exchange their spectrum sensing results $V_{i}$ with one hop neighbours. 
%Secondary users have the same transmission range on both licensed and control channel, and the communication links are reciprocal.
%
We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
Both primary users and CR users have fixed transmission ranges respectively, and all the channels are regarded to be identical in terms of signal propagation.
If a CR node locates within the transmission range of primary user $p$, that CR node is not allowed to use the channel $k(p)$.

We assume that in addition to the licensed channels, there is one dedicated control channel.
This control channel could be in ISM band or other reserved spectrum which is exclusively used for transmitting control messages.
Actually, the control messages involved in the clustering process can be transmitted on available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to any $i' \in \text{Nb}(i)$.
It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
$\text{Nb}(i)$ is simply defined as the set of nodes located within the transmission range of $i$.

@@

@@

\section{Clustering Algorithm: ROSS}
\label{ross}

In this section we introduce the distributed clustering scheme ROSS which leads to robust clusters against PU' influence.
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters on the basis of the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

@@

\item \textit{Social connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It is the number of common channels available to $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbours.
\end{itemize}
Individual connectivity degree $d_i$ and social connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where each node's connectivity vector is calculated and shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph and the connectivity vector $(d_i, g_i)$ for each node. The available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range. The number of common channels between two nodes is shown by the edge.}
	\label{fig1}
\end{figure}

@@

\subsubsection{Determining Cluster Heads and Form the Initial Clusters}
In this phase, each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than any neighbors except for those which have already become cluster heads (the appearance of cluster heads will be explained in Section~\ref{membershipClarification}), then node $i$ becomes clusters head.
%In other words, CR node $i$ becomes cluster head if $d_i > d_k, \forall k\in \texttt{Nb}_i\setminus CHs$ ($CHs$ donate the cluster heads existing in $\texttt{Nb}_i$).
If there is another CR node $j$ in its neighborhood which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\text{CHs}\cup i\}$, then the node out of $\{i, j\}$ with higher social connectivity degree becomes cluster head. The other nodes become a member of that cluster. 
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster's head.
%
The node which becomes cluster head broadcasts a message of its eligibility of being cluster head to notify its neighbours, and claims its neighbourhood as its cluster.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.

@@

\subsubsection{Guarantee the Availability of Common Control Channel}
%After deciding itself being cluster head, CR node broadcasts to notify its neighbours on the dedicated control channel, meanwhile, $i$'s initial cluster is formed immediately, which is $i$'s neighbourhood except for those nodes which have become cluster heads, \ie $C_i=(\text{Nb}_i\setminus \text{CHs})\cup i$.
After executing phase I of ROSS, there are some secondary users which become cluster heads.
The cluster head and its neighbourhood (except for the CHs) form a cluster.
It is possible that certain formed clusters don't own CCCs, then we solve this problem with the following method.

@@

In the following we illustrate the pressing necessity to control the cluster size when CRN becomes dense via both theoretical analysis and simulation.
Assuming the secondary and primary users are evenly distributed and primary users occupy the licensed channels randomly, then both CR nodes density and channel availability in the CRN can be seen to be spatially homogeneous.
%Based on Algorithm~\ref{alg0}, cluster heads are the CR nodes which possess the biggest individual connectivity degrees in their neighborhood respectively, and they are surrounded by CR nodes.
%In contrast, CR nodes residing on edge are unlikely to become cluster heads as their neighbourhoods are only half the nodes locating away from edge.
The formed clusters are the neighbourhoods of cluster heads, and the neighbourhood is decided by the transmission range and network density.
%
We consider a cluster $C(i)$ where $i$ is CH in a dense CRN. 
When we don't consider the CHs which could appear within $i$'s neighbourhood in the procedure of guaranteeing CCCs, according to Algorithm~\ref{alg0}, the nearest cluster heads could locate just outside node $i$'s transmission range.
An instance of this situation is shown in Figure~\ref{clusters_denseNetwork}.
%
In the figure, black dots represent cluster heads, the circles denotes the transmission ranges of cluster heads.
Cluster members are not shown in the figure.
%Circles represent the transmission range of cluster head, within which CR nodes are absorbed in cluster.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.3\linewidth]{clusters_denseNetwork_2.pdf}
  \caption{Clusters formation in extremely dense CRN. Black dots are cluster heads, other cluster members are not drawn.}
  \label{clusters_denseNetwork}
\end{figure}
Let $l$ be the length of side of simulation plan square, and $r$ be CR's transmission radius.
Based on the aforementioned analysis and geometry illustration as shown in Figure~\ref{clusters_denseNetwork}, we give an estimate on the maximum number of generated clusters, which is the product of the number of cluster heads in one row and that number in one line, $l/r * l/r = l^2/r^2$.
%Now we verify the estimation with simulation.
%We distribute CR and primary users randomly on a square plain, $r$=10 and $l$=50.
%Network density is increased by adding more CR users.
%We now have a look at how does the network density affect the cluster size when the transmission range is constant.
%This implicates when the cluster size is decided by the density of the network.
%As to SOC, the membership of one cluster is decided after a complex process, and the cluster size is roughly the same with one neighborhood.xxxx
%We can see from the example that although two neighbouring clusters can overlap greatly with each other, no cluster head will be covered by other clusters.
Given $r$=10 and $l$=50, the number of formed clusters is shown in Figure~\ref{number_clusters_scale}.
With the increase of CR users, network density increases linearly (the Y axis label is the average number of neighbours), and the number of formed clusters also increases and approaches to the the upper bound of 25 which complies with the estimation.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\linewidth]{number_clusters_upperBound.pdf}
  \caption{The correlation between the number of formed clusters and network density. Note that the number of neighbours denotes the network density. Simulation is run for 50 times and the confidence interval is 95\%.}
  \label{number_clusters_scale}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.5\linewidth]{figure2.pdf}
  \caption{Clusters formation after the first phase of ROSS. There are some nodes being debatable nodes, \ie belonging to more than one cluster.}
  \label{fig2}
\end{figure}

\subsection{Phase II - Membership Clarification}
\label{membershipClarification}
%\subsubsection*{Problem Description}
After applying phase I of ROSS to the example in Figure~\ref{fig1}, the resulted clusters are shown in Figure~\ref{fig2}.
We notice nodes $A, B, D$ are included in more than one cluster. 
We refer these nodes as \textit{debatable nodes} as their cluster affiliations are not clear, and the clusters which include the debatable node $i$ are called \textit{claiming clusters} of node $i$, and are denoted as $S_i$.  
Actually, debatable nodes extensively exist in CRN with larger scale.
Debatable nodes should be exclusively associated with only one cluster and be removed from the other claiming clusters, this procedure is called \textit{cluster membership clarification}.
We will introduce the solution for cluster membership clarification in the following.

@@

This procedure raises a concern that the debatable nodes may never stop changing their affiliations, as debatable nodes' choices seem to be dependent on each other, thus the infinite chain effect never ceases.
%  update their choices based on other debatable nodes' choices, and this process 
For example, assuming one debatable node $i$ locates in cluster $C\in S_i$, and $C$ has more than one debatable node except for $i$.
Assuming node $i$ makes decision to stay in the claiming cluster $C$, afterwards one another debatable nodes $j$ decides its affiliation, and there is $j\in C\in S_i$.
When $j$ leaves cluster $C$, which decrease cluster $C$'s cluster size, and could possibly trigger node $i$ to alter its previous decision to leave $C$, as $C$'s size is smaller now and leaving it may result in more increase of CCCs in $S_i$.
At this point of time, debatable node $j$ may rejoin cluster $C$ due to the changes in $S_j$, then both node $i$ and $j$ face the \textit{same}\footnote{Actually it is not totally same as before, as there are some new changes within $S_j$.} situation again.
%To maximize the increase of common channel in clusters in $S_j$, $j$ either stays in $C$, or leaves cluster $C$ and stay another cluster in $S_j$.
%$j$'s decision changes $C$'s membership, which will affect $i$'s decision.
%Assume $i$ makes decision before $j$ and chooses to stay in $C$, as which brings the most common channels in $i$' claiming clusters, and then it is $j$' turn to choose cluster.
%If $j$ leaves $C$, the smaller cluster $C$ will possibly make $i$ to leave it and join another cluster.
%and the choice of $j$ to stay in $C$ or not possibly changes $C$'s membership and  which potentially further triggers node $i$ to alter its previous decision. 
Thence, we need to answer this concern before implementing ROSS-DGA.
%, and if it converges, how good such a distributed scheme performs. 
In the following we show that the process of membership clarification can be formulated into a game, and a equilibrium is reached after a finite number of best response updates made by the debatable nodes.

@@

To formulate the debatable nodes' membership clarification into the desired congestion game, we observe this process from a different (or opposite) perspective. 
From the new perspective, the debatable nodes are regarded to be isolated and don't belong to any cluster, in other words, their claiming clusters become clusters which are beside them. 
Now for the debatable nodes, the previous problem of deciding which clusters to leave becomes a new problem that which cluster to join.
In the new problem, debatable node $i$ chooses one cluster $C$ out of $S_i$ to join if the decrease of CCCs in cluster $C$ is the smallest in $S_i$, and the decrease of CCCs in cluster $C$ is $\sum_{C\in S_i}\Delta\vert K(C) \vert=\sum_{C\in S_i}({\vert K(C) \vert-\vert K(C\cup i) \vert})$.
The interaction between the debatable nodes and the claiming clusters is shown in Figure~\ref{debatable_nodes_claiming_cluster}.
%The concern on convergence appears again as we have discussed in the previous subsubsection.
%We give proof on convergence under game theoretic framework.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.25\linewidth]{singletongame_png.png}
  \caption{Debatable nodes and claiming clusters}
  \label{debatable_nodes_claiming_cluster}
\end{figure}

@@

%\todo[inline]{check}
The centralized clustering scheme aims to form clusters with desired sizes, meanwhile the total number of CCCs of all clusters is maximized.
In the following, we refer this problem as \textit{centralized clustering}, and give the formal problem definition.

@@

The objective is a sum of two parts, the first part is the sum of products of cluster size and the corresponding number of CCCs, which is the sole metric adopted by the scheme SOC~\cite{Lazos09}.
The second part is the \textit{punishment} for choosing the clusters whose sizes are not $\delta$.
In fact, the second part is particularly designed to eliminates the drawbacks of SOC, \ie SOC produces a large number of singleton clusters and a few large clusters.
In practical computation, we minimize the opposite of the first part, and the punishment is a positive value.
%
The first constraint restricts each node $j$ to reside in exactly one cluster.
In the second constraint, $w_i$ is an auxiliary binary variable which denotes whether cluster $C_i$ is chosen by the solution, in particular, 
$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th legitimate cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th legitimate cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
The second constraint regulates when the $i$th legitimate cluster $C_i$ is chosen, the number of elements which are 1 in the $i$th row of the matrix $X$ is $|C_i|$.
%
Now we explain how does the mechanism of the punishment in the objective work. 
The parameter $p$ is defined as follows,
$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$} \\
\dots
\end{array} \right.
$$
where $\alpha_i>0$ and increases when $|C_i|$ diverges from $\delta$.
Because of $w_i$, any chosen cluster ($w=0$) brings certain \textit{punishment}.
When the chosen cluster's size is desired size $\delta$, the punishment is zero.
In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
In particular, when $w_i=0$ and $|C_i|=1$, the punishment is the most severe.
This design doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
Choice of $\alpha_i$ affects the resultant clusters.

@@

\item \textit{Cluster sizes.}
Specific clusters size is pursued in many applications due to energy preservation and the system design ~\cite{clustering_globecom11}.
We will present the distribution of CRs residing in the formed clusters with different sizes.
